\textbf{ Matching the forecast horizon with the relevant ecological processes }

### 2019-09-13 doc/manuscript.tex

Another approach to addressing the joint importance of spatial and time-series data for making ecological forecasts is joint spatiotemporal modeling...\comment{Ethan, what do you have in mind here? Probably something more than just working with spatiotemporal datasets and accounting for autocorrelation. Mevin has lots of papers with some sort of mechanistic modeling of spatiotemoral processes like dispersal and colonization. Should we say that the datasets needed to train this kind of model are rare, but that might be changing with more and more remote sensing data that is extensive in space and time? }

>>>>

Another potential approach to addressing the joint importance of spatial and time-series data for making ecological forecasts is to use models that incorporate both spatial and temporal information in data simultaneously. Spatiotemporal modeling approaches are increasingly common in ecology for patterns and processes that are of interest to forecsating (e.g., https://doi.org/10.1002/ecy.1643, https://doi.org/10.1890/0012-9658(2003)084[1382:HBMFPT]2.0.CO;2, https://doi.org/10.1111/geb.12666). However these models are rarely used in forecasting contexts. This is due in part to a combination of data limitation and computational challenges related to fitting and predicting from these models. Many data sources contain either spatial of temporal variation, but not both, and when spatiotemporal data is available it often involves irregular sampling, which creates challenges for modeling. Spatiotemporal models are also computationally intensive to fit and make predictions from, especially with large datasets (https://doi.org/10.1002/sta4.160). Fortunately thanks to large scale monitoring efforts like the National Ecological Observatory Network (citation), community science projects (e.g., eBird), and remote sensing large scale spatiotemporal data is increasingly available. In addition, new methods for spatiotemporal forecasting are being developed that address existing computational challenges (https://onlinelibrary.wiley.com/doi/full/10.1002/sta4.160) and access to high performance computing resources are increasingly common. As such, future ecological forecasting efforts should explore spatiotemporal approaches when data is available and assess whether these approaches produce improved predictions at intermediate time scales relative to spatial or temporal approaches alone.
